The Array of Long Baseline Antennas for Taking Radio Observations from the Sub-Antarctic

### 2020-10-27 albatros_paper.tex
started responses to referees
added some more comment responses

Very few experiments have surveyed the radio sky at $\lesssim 30$~MHz,
and here, we highlight only a subset.  At the very lowest frequencies,
the state of the art among ground-based measurements dates from the
1950s, when Reber and Ellis caught brief glimpses of the $2.1$~MHz sky
at $\sim 5$\degree\ resolution, using an array of
192~dipoles~\citep{1956JGR....61....1R}.  A few space-based missions
have also performed measurements at similarly low frequency ranges;
for example, the Radio Astronomy Explorer-2 operated at
\SIrange{0.025}{13}{\MHz} with $\sim 10$\degree\ resolution at
\SI{4.7}{\MHz}~\citep{1975A&A....40..365A}.  More recent ground-based
experiments include the Dominion Radio Astrophysical Observatory
surveys of the northern sky at 10~MHz and 22~MHz, with
$\sim2$\degree\ and 1.1--1.7\degree\ resolution,
respectively~\citep{1976MNRAS.177..601C, 1999A&AS..137....7R}.
Low-frequency imaging experiments that are currently operational or
under construction include the Ukranian T-shaped Radio
telescope~\citep{2016JAI.....541010Z}, with 0.5\degree\ resolution at
the center of its 8--32~MHz operating range, and the New extension in
Nan\c{c}ay upgrading LOFAR~\citep{7136773}, which operates at
10--85~MHz with 1.5\degree\ resolution in standalone mode, and
improved resolution at $>30$~MHz when combined with LOFAR.  Existing
measurements with finer resolution are mainly at higher frequencies;
for example, the Owens Valley Long Wavelength Array mapped the sky
with \SI{15}{\arcminute} resolution between 36.528~MHz and
73.152~MHz~\citep{2018AJ....156...32E}.  Although this experimental
list is not comprehensive, it does illustrate the dearth of
information about the $\lesssim 30$~MHz sky and the lack of high
resolution measurements at the lowest frequencies, particularly in the
southern hemisphere.  The first step toward laying the groundwork for
possible measurements of the dark ages is obtaining an improved map of
Galactic foregrounds at $\lesssim 30$~MHz.  In addition to serving as
a stepping stone for future cosmological constraints from the dark
ages, maps at these frequencies can also shed new light on Galactic
astrophysics.  Above $\sim30$~MHz, synchrotron emission from the
Galaxy follows a power law with a frequency dependence of $\sim
\nu^{-0.7}$.  At lower frequencies, synchrotron self-absorption
becomes non-negligible, and the spectrum transitions to $\nu^{+5/2}$
dependence.  Since self-absorption causes the optical depth along the
line of sight to depend on frequency, observations at $\lesssim
30$~MHz have the potential to probe the three-dimensional cosmic ray
structure of the Galaxy~\citep{2002ApJ...575..217P}.  Low-frequency
observations can also be used to image radio recombination lines
(RRLs), which provide a means for studying cool, low-density regions
of the interstellar medium (ISM).  These absorption lines arise from
Rydberg atoms and are highly sensitive to the surrounding environment.
The RRL spectrum and absorption profiles can therefore constrain the
detailed properties of the ISM regions in which the Rydberg atoms are
formed~\citep{2009NewAR..53..259G, 2007MNRAS.374..852S}.  Finally,
observations at $\lesssim30$~MHz may provide new views of Earth's
ionosphere, which absorbs and refracts at radio frequencies and
becomes completely opaque below the plasma cutoff frequency.  The
cutoff frequency and the levels of absorption and refraction are
time-varying and spatially dependent.  Experiments that image the
radio sky at multiple frequencies bracketing the plasma cutoff can
therefore simultaneously image the ionosphere and provide spectral
riometry data for space weather studies~\citep{2014GeoRL..41.5370K}.

@@

To assess the loss of sensitivity as the signal level increases, we
use \autoref{eqn:1bit_output} to derive the correlator efficiency
$\epsilon$, which is the ratio of measured signal-to-noise ratio (SNR)
to the ideal, infinite-precision SNR.  \autoref{fig:1bit_efficiency}
illustrates that the decrease in $\epsilon$ as a function of SNR is
relatively shallow.  When the SNR is near zero, $\epsilon$ approaches
$2/\pi \approx 0.64$, and if the signal level increases to two times
the  
noise level, $\epsilon$ decreases by only a factor of $\sim2$.  The
final \albatros\ array will have baselines that are long relative to
the observing wavelength, and the correlated component of the electric
field (which sets the relevant signal level) is likely a small
fraction of the total received power.  This qualitative argument can
by made by considering antennas in the $uv$~plane in the flat-sky
approximation.  Any individual antenna receives all power present in
the $uv$~plane (plus system noise), but the correlated contribution
comes only from the portion of the $uv$~plane that is encompassed by
the primary beam at the antenna's $uv$-space coordinate.  Assuming
that the sky can be described as a Gaussian random field, the maximum
correlated signal between two antennas is determined by the aperture
filling factor of that pair.  For the case of dipole antennas, the
filling factor is the square of the wavelength divided by the baseline
length.  Therefore, in the long-baseline limit, the correlated signal
is small in comparison to the total power received, and the correlator
efficiency is unlikely to decrease significantly from the ideal
$2/\pi$ value.

@@

\begin{figure}[t]
  \begin{center}
    \includegraphics[width=\linewidth]{Figures/albatros2_waterfalls.jpg}
    \caption{Spectra from a co-aligned polarization pair from the
      two-element \albatros\ pathfinder, shown as a function of
      frequency and time.  The auto- and cross-spectra are shown in
      the top and bottom rows, respectively.  The spectrum magnitudes
      are in uncalibrated ADC units on a logarithmic scale, and the
      cross-spectrum is about two orders of magnitude fainter than the
      autospectra.  Repeatable structure from the Galaxy is visible
      over the three-day time scale in all plots.}
    \label{Fig:albatros2_waterfalls}
  \end{center}
\end{figure}

>>>>

Very few experiments have surveyed the radio sky at $\lesssim 30$~MHz,
and here, we highlight only a subset.  At the very lowest frequencies,
the state of the art among ground-based measurements dates from the
1950s, when Reber and Ellis caught brief glimpses of the $2.1$~MHz sky
at $\sim 5$\degree\ resolution, using an array of
192~dipoles~\citep{1956JGR....61....1R}.  A few space-based missions
have also performed measurements at similarly low frequency ranges;
for example, the Radio Astronomy Explorer-2 operated at
\SIrange{0.025}{13}{\MHz} with $\sim 10$\degree\ resolution at
\SI{4.7}{\MHz}~\citep{1975A&A....40..365A}.  More recent ground-based
experiments include the Dominion Radio Astrophysical Observatory
surveys of the northern sky at 10~MHz and 22~MHz, with
$\sim2$\degree\ and 1.1--1.7\degree\ resolution,
respectively~\citep{1976MNRAS.177..601C, 1999A&AS..137....7R}.
Low-frequency imaging experiments that are currently operational or
under construction include the Ukranian T-shaped Radio
telescope~\citep{2016JAI.....541010Z}, with 0.5\degree\ resolution at
the center of its 8--32~MHz operating range, and the New extension in
Nan\c{c}ay upgrading LOFAR~\citep{7136773}, which operates at
10--85~MHz with 1.5\degree\ resolution in standalone mode, and
improved resolution at $>30$~MHz when combined with LOFAR.  Existing
measurements with finer resolution are mainly at higher frequencies;
for example, the Owens Valley Long Wavelength Array mapped the sky
with \SI{15}{\arcminute} resolution between 36.528~MHz and
73.152~MHz~\citep{2018AJ....156...32E}.  Although this experimental
list is not comprehensive, it does illustrate the dearth of
information about the $\lesssim 30$~MHz sky and the lack of high
resolution measurements at the lowest frequencies, particularly in the
southern hemisphere.

The first step toward laying the groundwork for
possible measurements of the dark ages is obtaining an improved map of
Galactic foregrounds at $\lesssim 30$~MHz.  In addition to serving as
a stepping stone for future cosmological constraints from the dark
ages, maps at these frequencies can also shed new light on Galactic
astrophysics.  Above $\sim30$~MHz, synchrotron emission from the
Galaxy follows a power law with a frequency dependence of $\sim
\nu^{-0.7}$.  At lower frequencies, synchrotron self-absorption
becomes non-negligible, and the spectrum transitions to $\nu^{+5/2}$
dependence.  Since self-absorption causes the optical depth along the
line of sight to depend on frequency, observations at $\lesssim
30$~MHz have the potential to probe the three-dimensional cosmic ray
structure of the Galaxy~\citep{2002ApJ...575..217P}.  Low-frequency
observations can also be used to image radio recombination lines
(RRLs), which provide a means for studying cool, low-density regions
of the interstellar medium (ISM).  These absorption lines arise from
Rydberg atoms and are highly sensitive to the surrounding environment.
The RRL spectrum and absorption profiles can therefore constrain the
detailed properties of the ISM regions in which the Rydberg atoms are
formed~\citep{2009NewAR..53..259G, 2007MNRAS.374..852S}.  Finally,
observations at $\lesssim30$~MHz may provide new views of Earth's
ionosphere, which absorbs and refracts at radio frequencies and
becomes completely opaque below the plasma cutoff frequency.  The
cutoff frequency and the levels of absorption and refraction are
time-varying and spatially dependent.  Experiments that image the
radio sky at multiple frequencies bracketing the plasma cutoff can
therefore simultaneously image the ionosphere and provide spectral
riometry data for space weather studies~\citep{2014GeoRL..41.5370K}.

@@

To assess the loss of sensitivity as the signal level increases, we
use \autoref{eqn:1bit_output} to derive the correlator efficiency
$\epsilon$, which is the ratio of measured signal-to-noise ratio (SNR)
to the ideal, infinite-precision SNR.  \autoref{fig:1bit_efficiency}
illustrates that the decrease in $\epsilon$ as a function of SNR is
relatively shallow.  When the SNR is near zero, $\epsilon$ approaches
$2/\pi \approx 0.64$, and if the signal level increases to two times
the  
noise level, $\epsilon$ decreases by only a factor of $\sim2$.  The
final \albatros\ array will have baselines that are long relative to
the observing wavelength, and the correlated component of the electric
field (which sets the relevant signal level) is likely a small
fraction of the total received power.  This qualitative argument can
by made by considering antennas in the $uv$~plane in the flat-sky
approximation.  Any individual antenna receives all power present in
the $uv$~plane (plus system noise), but the correlated contribution
comes only from the portion of the $uv$~plane that is encompassed by
the primary beam at the antenna's $uv$-space coordinate.  Assuming
that the sky can be described as a Gaussian random field, the maximum
correlated signal between two antennas is determined by the aperture
filling factor of that pair.  For the case of dipole antennas, the
filling factor is the square of the wavelength divided by the baseline
length.  Therefore, in the long-baseline limit, the correlated signal
is small in comparison to the total power received, and the correlator
efficiency is unlikely to decrease significantly from the ideal
$2/\pi$ value.  This remains true even when the brightest radio
sources in the sky are in the beam (the sun is irrelevant since
ionospheric effects restrict \albatros\ to night-time observing).
Centaurus A \citep{culgoora1} and M1 \citep{Braude1969} only affect
the total power expected from the Global Sky Model \citep{gsm2008} at
the $\sim$10\% level.  Additionally, as will be shown later in this
work, preliminary ALBATROS data show that visibilities are only a few
percent of the total power amplitude, consistent with our expectation
that diffuse emission dominates even the brightest sources given the
$\sim$steradian solid angle of the primary beam.

@@

\begin{figure}[t]
  \begin{center}
    \includegraphics[width=\linewidth]{Figures/albatros2_waterfalls.jpg}
    \caption{Spectra from a co-aligned polarization pair from the
      two-element \albatros\ pathfinder, shown as a function of
      frequency and time.  The auto- and cross-spectra are shown in
      the top and bottom rows, respectively.  The spectrum magnitudes
      are in uncalibrated ADC units on a logarithmic scale, and the
      cross-spectrum is about two orders of magnitude fainter than the
      autospectra (.  Repeatable structure from the Galaxy is visible
      over the three-day time scale in all plots.}
    \label{Fig:albatros2_waterfalls}
  \end{center}
\end{figure}
